An Experimental Evaluation of Hybrid Vector Search (EA\&B)

### 2025-04-28 VLDB2026/main.tex
update main

		\centering
		%	 \setlength{\abovecaptionskip}{0cm}
		%	\setlength{\belowcaptionskip}{0.1cm}
		\footnotesize{
			\begin{minipage}{\linewidth}
				\vspace{0.1cm}
				\textsuperscript{A} Post-filtering, 
				\textsuperscript{B} Pre-filtering, 
				\textsuperscript{C} Simultaneous filtering, 
				\textsuperscript{Y} Support, 
				\textsuperscript{N} Unsupport. 
				\textit{Base} indicates that the hybrid query algorithm is an improvement based on a specific ANN search algorithm. 
				\textit{AND} indicates whether the algorithm supports attribute filtering with AND operations. 
				\textit{OR} indicates whether the algorithm supports attribute filtering with OR operations. 
				\textit{Flexible Attributes} indicates whether the number of query attributes can differ from the number of attributes used during index construction. 
				\textit{Complex Boolean} refers to whether the algorithm supports advanced Boolean logic beyond just AND and OR, such as NOT or combinations like (A AND B) OR (C AND NOT D). 
				\textit{Dynamic} indicates whether the algorithm supports the operation of dynamically inserting data points.
		\end{minipage}}
		\vspace{-0.5cm}
	\end{table*}

@@

	\noindent\textbf{\underline{Filtered-DiskANN \cite{Filtered-diskann}.}}  
	\sout{The effectiveness of NHQ may degrade when the number of attributes changes, as it affects the fusion distance computation. Filtered-DiskANN addresses this limitation.}
	Filtered-DiskANN also supports simultaneous filtering. Built upon the Vamana \cite{diskann} graph-based ANN index, it incorporates attribute information directly into the graph during index construction. This integration ensures that the index reflects both vector similarity and attribute constraints. Moreover, Filtered-DiskANN supports SSD-based storage, improving scalability.
	Filtered-DiskANN proposes 2 indexes:  
	1) \textbf{FilteredVamana}, which incrementally builds the graph index by inserting data points and dynamically adding edges, allowing adaptive expansion.  
	2) \textbf{StitchedVamana}, which adopts a batch construction strategy—constructing separate Vamana subgraphs for each attribute, followed by merging and edge pruning.
	A key limitation of Filtered-DiskANN is that while it supports both single-attribute and multi-attribute filtering, it only supports the Boolean OR logic for multi-attribute filtering, lacking support for Boolean AND logic.

	\noindent\textbf{\underline{CAPS \cite{CAPS}.}} 
	\sout{Different from the above graph-based simultaneous filtering algorithms,}CAPS is the first simultaneous filtering algorithm based on spatial partitioning.
	%As graph-based methods, NHQ and Filtered-DiskANN often result in large index sizes. CAPS addresses this issue by being the first simultaneous filtering algorithm based on spatial partitioning, significantly reducing index size.
	CAPS introduces a hierarchical sub-partitioning algorithm inspired by Huffman trees, termed the Attribute Frequency Tree (AFT), to overcome the coarse granularity of traditional IVF-based methods. CAPS adopts a two-level partitioning strategy:  
	1) The first level clusters vectors based on similarity using K-Means or learning-based methods such as BLISS. 2) Within each cluster, AFT partitions data further based on attribute frequencies, enabling finer-grained indexing and improving query efficiency.

@@

	\noindent\textbf{\underline{iRange \cite{iRangeGraph}.}}  
	To address the query efficiency degradation caused by the graph compression in SeRF, iRange adopts a more flexible strategy. Rather than prebuilding indexes for all possible ranges, it dynamically assembles a query-specific subgraph at runtime. iRange partitions the dataset into intervals based on attribute values and independently constructs a local graph for each interval, storing only edge information. During the query process, the relevant local graphs overlapping with the query range are merged into a temporary search graph, and a pruning strategy is applied to enable efficient search.

@@

	%4.2.1
	\subsubsection{Time and Space overhead of index construction}
	% \subsubsection{\textbf{Attribute Filtering}}
	\begin{figure}[t]
		\centering
		\setlength{\abovecaptionskip}{0.3cm}
		\setlength{\belowcaptionskip}{-0.15cm}

@@

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0cm}
		\setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=0.96\textwidth]{figures/exp/exp_1_1_SingleLabel_1thread.pdf}
		\caption{Single-Attribute Building and Query (single thread) }
		\label{fig:exp_1_1_SingleLabel_1thread}
	\end{figure*}

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0cm}
		\setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=0.96\textwidth]{figures/exp/exp_1_2_SingleLabel_16thread.pdf}
		\caption{Single-Attribute Building and Query (16 threads)}
		\label{fig:exp_1_2_SingleLabel_16thread}
	\end{figure*}

@@

	\begin{figure}[t]
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.1cm}
		% 上面的图例，居中并可通过 hspace 调整左右位置
		\hspace*{15pt} % 可调整的参数，负值左移，正值右移，例如 -20pt 或 20pt
		\includegraphics[width=0.95\columnwidth]{figures/exp/exp_2_legend.pdf} % 图例图片路径

@@

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0.0cm}
		\setlength{\belowcaptionskip}{-0.4cm}
		\includegraphics[width=0.95\textwidth]{figures/exp/exp_3_1.pdf}
		\caption{Effect of attribute distribution on query performance (single thread)}
		\label{fig:exp_3_1}
	\end{figure*}

@@

		% 第一行：单独一张大图
		\begin{subfigure}{\textwidth}
			\centering
			\setlength{\abovecaptionskip}{0.05cm}
			\setlength{\belowcaptionskip}{0.05cm}
			\includegraphics[width=\textwidth]{figures/exp/exp_5_2_1.pdf}
			\caption{Algorithms optimized for low AS levels.}
			\label{fig:exp_5_2_1}
		\end{subfigure}

@@

		% 第二行：两张子图，左侧宽度是右侧的两倍
		\begin{subfigure}{0.66\textwidth} % 2/3 宽度
			\centering
			\setlength{\abovecaptionskip}{0.05cm}
			\setlength{\belowcaptionskip}{0.05cm}
			\includegraphics[width=\textwidth]{figures/exp/exp_5_2_2.pdf}
			\caption{Algorithms optimized for high AS levels.}
			\label{fig:exp_5_2_2}
		\end{subfigure}
		\hfill % 增加水平间距
		\begin{subfigure}{0.33\textwidth} % 1/3 宽度
			\centering
			\setlength{\abovecaptionskip}{0.05cm}
			\setlength{\belowcaptionskip}{0.05cm}
			\includegraphics[width=\textwidth]{figures/exp/exp_5_2_3.pdf}
			\caption{Algorithms insensitive to AS level.}
			\label{fig:exp_5_2_3}
		\end{subfigure}

		\setlength{\abovecaptionskip}{0cm}
		\setlength{\belowcaptionskip}{-0.5cm}
		\caption{Different AS under the same algorithm}
		\label{fig:exp_5_2_combined}
	\end{figure*}

@@

	\begin{figure*}
		\centering
		    \setlength{\abovecaptionskip}{-0.1cm}
		    \setlength{\belowcaptionskip}{-0.4cm}
		\includegraphics[width=0.97\textwidth]{figures/exp/exp_6_1.pdf}
		\caption{Effect of dataset on Query}
		\label{fig:exp_6_1}
	\end{figure*}

@@

		\setlength{\abovecaptionskip}{0.02cm}
		\setlength{\belowcaptionskip}{-0.1cm}
		\hspace*{11pt}
		% 放大的图例（占满单栏）
		\includegraphics[width=0.93\columnwidth]{figures/indexData/rangeFilter_legend_only.pdf}

>>>>

		\centering
		%	 \setlength{\abovecaptionskip}{0cm}
		%	\setlength{\belowcaptionskip}{0.1cm}
		\footnotesize{
			\begin{minipage}{\linewidth}
				\vspace{0.1cm}
				\textsuperscript{A} Post-filtering, 
				\textsuperscript{B} Pre-filtering, 
				\textsuperscript{C} Simultaneous filtering, 
				\textsuperscript{Y} Support, 
				\textsuperscript{N} Unsupport. 
				\textit{Base} indicates that the hybrid query algorithm is an improvement based on a specific ANN search algorithm. 
				\textit{AND} indicates whether the algorithm supports attribute filtering with AND operations. 
				\textit{OR} indicates whether the algorithm supports attribute filtering with OR operations. 
				\textit{Flexible Attributes} indicates whether the number of query attributes can differ from the number of attributes used during index construction. 
				\textit{Complex Boolean} refers to whether the algorithm supports advanced Boolean logic beyond just AND and OR, such as NOT or combinations like (A AND B) OR (C AND NOT D). 
				\textit{Dynamic Insert} indicates whether the algorithm supports the operation of dynamically inserting data points.
				\textit{Multi Thread} indicates whether the algorithm supports multi-threaded search.
		\end{minipage}}
		\vspace{-0.5cm}
	\end{table*}

@@

	\noindent\textbf{\underline{Filtered-DiskANN \cite{Filtered-diskann}.}}  
	%\sout{The effectiveness of NHQ may degrade when the number of attributes changes, as it affects the fusion distance computation. Filtered-DiskANN addresses this limitation.}
	Filtered-DiskANN also supports simultaneous filtering. Built upon the Vamana \cite{diskann} graph-based ANN index, it incorporates attribute information directly into the graph during index construction. This integration ensures that the index reflects both vector similarity and attribute constraints. Moreover, Filtered-DiskANN supports SSD-based storage, improving scalability.
	Filtered-DiskANN proposes 2 indexes:  
	1) \textbf{FilteredVamana}, which incrementally builds the graph index by inserting data points and dynamically adding edges, allowing adaptive expansion.  
	2) \textbf{StitchedVamana}, which adopts a batch construction strategy—constructing separate Vamana subgraphs for each attribute, followed by merging and edge pruning.
%	A key limitation of Filtered-DiskANN is that only supports the Boolean OR logic for multi-attribute filtering, lacking support for Boolean AND logic.

	\noindent\textbf{\underline{CAPS \cite{CAPS}.}} 
	%\sout{Different from the above graph-based simultaneous filtering algorithms,}CAPS is the first simultaneous filtering algorithm based on spatial partitioning.
	%As graph-based methods, NHQ and Filtered-DiskANN often result in large index sizes. CAPS addresses this issue by being the first simultaneous filtering algorithm based on spatial partitioning, significantly reducing index size.
	CAPS introduces a hierarchical sub-partitioning algorithm inspired by Huffman trees, termed the Attribute Frequency Tree (AFT), to overcome the coarse granularity of traditional IVF-based methods. CAPS adopts a two-level partitioning strategy:  
	1) The first level clusters vectors based on similarity using K-Means or learning-based methods such as BLISS. 2) Within each cluster, AFT partitions data further based on attribute frequencies, enabling finer-grained indexing and improving query efficiency.

@@

	\noindent\textbf{\underline{iRange \cite{iRangeGraph}.}}  
	To address the query efficiency degradation caused by the graph compression in SeRF, iRange adopts a more flexible strategy.
	%Rather than prebuilding indexes for all possible ranges, it dynamically assembles a query-specific subgraph at runtime.
	iRange partitions the dataset into intervals based on attribute values and independently constructs a local graph for each interval, storing only edge information. During the query process, the relevant local graphs overlapping with the query range are merged into a temporary search graph, and a pruning strategy is applied to enable efficient search.

@@

	%4.2.1
	\subsubsection{Time and Space overhead of index construction}
	% \subsubsection{\textbf{Attribute Filtering}}
	\begin{figure}[t]
		\centering
		\setlength{\abovecaptionskip}{0.3cm}
		\setlength{\belowcaptionskip}{-0.0cm}

@@

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0cm}
		\setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=0.95\textwidth]{figures/exp/exp_1_1_SingleLabel_1thread.pdf}
		\caption{Single-Attribute Building and Query (single thread) }
		\label{fig:exp_1_1_SingleLabel_1thread}
	\end{figure*}

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0cm}
		\setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=0.95\textwidth]{figures/exp/exp_1_2_SingleLabel_16thread.pdf}
		\caption{Single-Attribute Building and Query (16 threads)}
		\label{fig:exp_1_2_SingleLabel_16thread}
	\end{figure*}

@@

	\begin{figure}[t]
		\centering
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.1cm}
		% 上面的图例，居中并可通过 hspace 调整左右位置
		\hspace*{15pt} % 可调整的参数，负值左移，正值右移，例如 -20pt 或 20pt
		\includegraphics[width=0.85\columnwidth]{figures/exp/exp_2_legend.pdf} % 图例图片路径

@@

	\begin{figure*}
		\centering
		\setlength{\abovecaptionskip}{0.0cm}
		\setlength{\belowcaptionskip}{-0.5cm}
		\includegraphics[width=0.95\textwidth]{figures/exp/exp_3_1.pdf}
		\caption{Effect of attribute distribution on query performance (single thread)}
		\label{fig:exp_3_1}
	\end{figure*}

@@

		% 第一行：单独一张大图
		\begin{subfigure}{\textwidth}
			\centering
			\setlength{\abovecaptionskip}{0.05cm}
			\setlength{\belowcaptionskip}{0.05cm}
			\includegraphics[width=0.95\textwidth]{figures/exp/exp_5_2_1.pdf}
			\caption{Algorithms optimized for low AS levels.}
			\label{fig:exp_5_2_1}
		\end{subfigure}

@@

		% 第二行：两张子图，左侧宽度是右侧的两倍
		\begin{subfigure}{0.66\textwidth} % 2/3 宽度
			\centering
			\setlength{\abovecaptionskip}{0.05cm}
			\setlength{\belowcaptionskip}{0.05cm}
			\includegraphics[width=0.92\textwidth]{figures/exp/exp_5_2_2.pdf}
			\caption{Algorithms optimized for high AS levels.}
			\label{fig:exp_5_2_2}
		\end{subfigure}
		\hfill % 增加水平间距
		\begin{subfigure}{0.33\textwidth} % 1/3 宽度
			\centering
			\setlength{\abovecaptionskip}{0.05cm}
			\setlength{\belowcaptionskip}{0.05cm}
			\includegraphics[width=0.92\textwidth]{figures/exp/exp_5_2_3.pdf}
			\caption{Algorithms insensitive to AS level.}
			\label{fig:exp_5_2_3}
		\end{subfigure}

		\setlength{\abovecaptionskip}{0cm}
		\setlength{\belowcaptionskip}{-0.4cm}
		\caption{Different AS under the same algorithm}
		\label{fig:exp_5_2_combined}
	\end{figure*}

@@

	\begin{figure*}
		\centering
		    \setlength{\abovecaptionskip}{-0.1cm}
		    \setlength{\belowcaptionskip}{-0.4cm}
		\includegraphics[width=0.95\textwidth]{figures/exp/exp_6_1.pdf}
		\caption{Effect of dataset on Query}
		\label{fig:exp_6_1}
	\end{figure*}

@@

		\setlength{\abovecaptionskip}{0.02cm}
		\setlength{\belowcaptionskip}{-0.1cm}
		\hspace*{11pt}
		% 放大的图例（占满单栏）
		\includegraphics[width=0.90\columnwidth]{figures/indexData/rangeFilter_legend_only.pdf}
